Open-source tools in R for landscape ecology

### 2021-04-25 Comments_Reviewer.Rmd
Updating Fig 1 and some more minor reviewer comments

**Reference:** Rayfield B, MJ Fortin, A Fall. 2011. Connectivity for conservation: a framework to classify network measures. Ecology 92(4): 847-858.

@@

@@

@@

>>>>

*Reference:* Rayfield B, MJ Fortin, A Fall. 2011. Connectivity for conservation: a framework to classify network measures. Ecology 92(4): 847-858.

@@

**Reply**: Thank you for the general encouraging words about the manuscript. 
We agree that a definition of landscape ecology is not needed given the target audience of the journal. 
Thus, we shortened the introduction. 
\textcolor{red}{Maybe move section of powerful R to introduction? But, comprehensive comparison with other software outside of scope} \vskip 0.5cm

@@

**Reply**: All authors of the manuscript are experienced R users with many years of experience using the programming language. 
Additionally, all authors even contributed several packages to the community themselves. 
Given the different background of all authors, collectively probably most packages were used by at least one of the authors, but probably not all. 
Nevertheless, we also believe that this is not necessarily important for this review. 
Furthermore, we agree with the reviewer that more specific recommendation which package to use could be useful, however, we also think that this is very difficult. 
Reasons for this include, besides others, personal preferences of the users or analysis specific circumstances. 
Lastly, the review should be as neutral as possible without favoring one package over another because pros and cons of package differ between users and all were developed with a specific idea in mind.
\textcolor{red}{Refer to new how to get started box} \vskip 0.5cm

@@

**Reply**: We agree with the reviewer that Figure 1 of Joo et al. places the packages nicely into a context. We tried to achieve something similar with our Figure 1 as well, but obviously failed to achieve this goal. Thus, we tried to improve Figure 1 in the revised version of the manuscript to highlight better how the packages could be placed in a workflow. Also, we added an explanation that only packages were included that were available on CRAN at the time of preparation of the manuscript, which ensures at least that packages are maintained to fulfill all technical requirements by CRAN. \vskip 0.5cm

### 2021-04-25 paper/paper.Rmd
Updating Fig 1 and some more minor reviewer comments

Even though many other open-source tools \cite{QGISDevelopmentTeam2016,GRASSDevelopmentTeam2017,Porta2017} and suitable programming languages (e.g. Python) for landscape ecology exist, in this review we focus on software implemented in the *R* programming language.
\textcolor{blue}{We acknowledges that there are reasons and circumstances to also use other software tools and progamming languages, however, a comprehensive discussion of pro and cons is outside the scope of this manuscript.}
For more general overviews of open-source software for \textcolor{blue}{landscape ecology studies} see \cite{Jolma2008,Steiniger2009,Steiniger2009a,Istvan2012}.
The *R* programming language can be a very powerful tool. In addition to handling spatial data, other analytical tasks such as statistical modeling, creation of publication-ready figures, and even complete reports can be done within the *R* environment (Fig. \@ref(fig:fig-workflow)).
The growing body of *R* packages related to spatial data processing and analysis results in a high capability of this language for landscape ecology.

Thus, in the first part of this article, we present an overview of existing *R* packages for landscape ecology and closely related fields (Table \@ref(tab:packages)).
In the second part, we present a survey in which we asked the community how they currently use the *R* programming language and to identify topics for which *R* packages are presently missing for landscape ecology.

@@

(ref:fig1cap) Exemplary workflow of spatial data analyses for landscape ecology using the *R* programming language. For all major tasks, some example *R* packages are listed.
```{r fig-workflow, echo = FALSE, out.width="\\linewidth", fig.cap = "(ref:fig1cap)"}
include_graphics("data/Figure_1.pdf")
```

>>>>

Even though many other open-source tools \cite{QGISDevelopmentTeam2016,GRASSDevelopmentTeam2017,Porta2017} and suitable programming languages (e.g. Python) for landscape ecology exist, in this review we focus on software implemented in the *R* programming language.
\textcolor{blue}{We acknowledges that there are reasons and circumstances to also use other software tools and progamming languages, however, a comprehensive discussion of pro and cons is outside the scope of this manuscript.}
For more general overviews of open-source software for \textcolor{blue}{landscape ecology studies} see \cite{Jolma2008,Steiniger2009,Steiniger2009a,Istvan2012}.

The *R* programming language can be a very powerful tool. In addition to handling spatial data, other analytical tasks such as statistical modeling, creation of publication-ready figures, and even complete reports can be done within the *R* environment (Fig. \@ref(fig:fig-workflow)).
The growing body of *R* packages related to spatial data processing and analysis results in a high capability of this language for landscape ecology.
Thus, in the first part of this article, we present an overview of existing *R* packages for landscape ecology and closely related fields (Table \@ref(tab:packages)).
\textcolor{blue}{We included only packages that were available on the official *R* package archive network CRAN at the time the manuscript was prepared. 
This ensures that all included packages were at least maintained to fulfill CRAN's technical quality standards.}
In the second part, we present a survey in which we asked the community how they currently use the *R* programming language and to identify topics for which *R* packages are presently missing for landscape ecology.

@@

(ref:fig1cap) Exemplary workflow of spatial data analyses for landscape ecology using the *R* programming language. For all major tasks, some example *R* packages are listed. For more packages related to the specific task, see Table \@ref(tab:packages).
```{r fig-workflow, echo = FALSE, out.width="\\linewidth", fig.cap = "(ref:fig1cap)"}
include_graphics("data/Figure_1.pdf")
```

### 2021-04-25 paper/paper.tex
Updating Fig 1 and some more minor reviewer comments

Even though many other open-source tools \cite{QGISDevelopmentTeam2016,GRASSDevelopmentTeam2017,Porta2017} and suitable programming languages (e.g.~Python) for landscape ecology exist, in this review we focus on software implemented in the \emph{R} programming language.
\textcolor{blue}{We acknowledges that there are reasons and circumstances to also use other software tools and progamming languages, however, a comprehensive discussion of pro and cons is outside the scope of this manuscript.}
For more general overviews of open-source software for \textcolor{blue}{landscape ecology studies} see \cite{Jolma2008,Steiniger2009,Steiniger2009a,Istvan2012}.
The \emph{R} programming language can be a very powerful tool. In addition to handling spatial data, other analytical tasks such as statistical modeling, creation of publication-ready figures, and even complete reports can be done within the \emph{R} environment (Fig. \ref{fig:fig-workflow}).
The growing body of \emph{R} packages related to spatial data processing and analysis results in a high capability of this language for landscape ecology.

Thus, in the first part of this article, we present an overview of existing \emph{R} packages for landscape ecology and closely related fields (Table \ref{tab:packages}).
In the second part, we present a survey in which we asked the community how they currently use the \emph{R} programming language and to identify topics for which \emph{R} packages are presently missing for landscape ecology.

@@

\begin{figure}
\includegraphics[width=\linewidth]{data/Figure_1} \caption{Exemplary workflow of spatial data analyses for landscape ecology using the \emph{R} programming language. For all major tasks, some example \emph{R} packages are listed.}\label{fig:fig-workflow}
\end{figure}

>>>>

Even though many other open-source tools \cite{QGISDevelopmentTeam2016,GRASSDevelopmentTeam2017,Porta2017} and suitable programming languages (e.g.~Python) for landscape ecology exist, in this review we focus on software implemented in the \emph{R} programming language.
\textcolor{blue}{We acknowledges that there are reasons and circumstances to also use other software tools and progamming languages, however, a comprehensive discussion of pro and cons is outside the scope of this manuscript.}
For more general overviews of open-source software for \textcolor{blue}{landscape ecology studies} see \cite{Jolma2008,Steiniger2009,Steiniger2009a,Istvan2012}.

The \emph{R} programming language can be a very powerful tool. In addition to handling spatial data, other analytical tasks such as statistical modeling, creation of publication-ready figures, and even complete reports can be done within the \emph{R} environment (Fig. \ref{fig:fig-workflow}).
The growing body of \emph{R} packages related to spatial data processing and analysis results in a high capability of this language for landscape ecology.
Thus, in the first part of this article, we present an overview of existing \emph{R} packages for landscape ecology and closely related fields (Table \ref{tab:packages}).
\textcolor{blue}{We included only packages that were available on the official *R* package archive network CRAN at the time the manuscript was prepared. 
This ensures that all included packages were at least maintained to fulfill CRAN's technical quality standards.}
In the second part, we present a survey in which we asked the community how they currently use the \emph{R} programming language and to identify topics for which \emph{R} packages are presently missing for landscape ecology.

@@

\begin{figure}
\includegraphics[width=\linewidth]{data/Figure_1} \caption{Exemplary workflow of spatial data analyses for landscape ecology using the \emph{R} programming language. For all major tasks, some example \emph{R} packages are listed. For more packages related to the specific task, see Table \ref{tab:packages}.}\label{fig:fig-workflow}
\end{figure}
